Audio Interface of an Active Perception Navigation Aid for People with Visual Impairments

### 2019-07-09 paper.tex
Added new elevation and hardware descriptions, as well as more references

\maketitle

@@

In recent years, governments have spearheaded numerous initiatives to support people with disabilities and enable them to play a more active role in modern society.
The UK's Royal National Institute of Blind People (RNIB) for example has prioritised improving access to everyday services and products, such as public transport and mobile apps~\cite{rnib-objectives}.
Improvements in modern computing have made it possible for new and innovative solutions to these problems to come to the fore.
In particular, researchers in the active vision field have have made much progress in enabling machines to autonomously manipulate cameras to gather information about an environment for mapping and object and object finding tasks~\cite{bajcsy2018revisiting}.
There is, however, a significant research question about whether techniques from active machine vision can be applied to humans, i.e.\ can a machine identify a point of interest in a scene and direct a human, instead of an electronic servo, to focus on that point?
If this can be done, it would be beneficial to people with visual impairments and will augment their ability to search for an arbitrary point or object of interest and identify an unknown scene.

@@

%For this work, we investigate implementing such an interface that uses audio signals to transmit guidance instructions by exploiting the human auditory and psychoacoustic systems. 
%Initially the interface is implemented in 2D only (pan and tilt dimensions) to provide the user with audio instructions on where to point the device and our approach was validated through a set of simple pointing tasks.

\begin{figure}
  \centering
  \subfloat[]{\label{fig:tango-headphone}\includegraphics[width=0.4\columnwidth]{figures/tango_headphone.png}}
  %\includegraphics[width=0.5\columnwidth]{figures/tango_headphone.png}}
  ~
  \subfloat[]{\label{fig:participant}\includegraphics[width=0.25\columnwidth]{figures/vi_participant.pdf}}
  \caption{Pictures of the Tango device and bone-conducting headphones the system is based on (left) and both in use in an experiment (right).}\label{fig:tango-headphone}
\end{figure}

@@

\section{Interface Implementation}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.3\columnwidth]{figures/camera_coordinate.pdf}
  \caption{A diagram showing the device's frame of reference and the target's pan and tilt angles.}\label{fig:cam-coords}
\end{figure}

%The audio signal used for this work is a simple \SI{10}{\hertz} sinusoidal wave that is easy to manipulate and analyse.
%People have trouble localising a pure sine wave given its lack of spectral features

@@

The human audition system uses binaural comparison cues, such as the inter-aural time and level differences (ITD and ILD respectively), to localise a sound source on the horizontal plane.
The ITD is the perceived time delay between the signal reaching both ears, while the ILD is the perceived volume difference in the signal.
For example, a sound that comes from the individual's right, will hit the right ear first with a slightly higher volume.

In this work, a pure sinusoidal wave oscillating at \SI{10}{\hertz} was used. 
People typically have trouble localising a pure tone devoid of a rich spectral signature.
However, the ITD and ILD is independent of the tone's spectral make-up and the elevation angle is given through a different mechanism.
A pure sine wave is therefore sufficient to convey the target's pan angle.
To transform the signal, we used OpenAL's default HRTF, based on the MIT's KEMAR dataset\cite{hiebert2005openal,gardner1995hrtf}, that uses the user and targets' positions as input and outputs a transformed audio signal.

@@

>>>>

\maketitle
\setcounter{footnote}{0}

@@

In recent years, governments have spearheaded numerous initiatives to support people with disabilities and enable them to play a more active role in modern society.
The UK's Royal National Institute of Blind People (RNIB) for example has prioritised improving access to everyday services and products, such as public transport and mobile apps~\cite{rnib2016uk}.
Improvements in modern computing have made it possible for new and innovative solutions to these problems to come to the fore.
In particular, researchers in the active vision field have have made much progress in enabling machines to autonomously manipulate cameras to gather information about an environment for mapping and object and object finding tasks~\cite{bajcsy2018revisiting}.
There is, however, a significant research question about whether techniques from active machine vision can be applied to humans, i.e.\ can a machine identify a point of interest in a scene and direct a human, instead of an electronic servo, to focus on that point?
If this can be done, it would be beneficial to people with visual impairments and will augment their ability to search for an arbitrary point or object of interest and identify an unknown scene.

@@

\begin{figure}
  \centering
  \subfloat[]{\label{fig:tango-headphone}\includegraphics[width=0.4\columnwidth]{figures/tango_headphone.png}}
~
  \subfloat[]{\label{fig:participant}\includegraphics[width=0.25\columnwidth]{figures/vi_participant.pdf}}
  \caption{Pictures of the Tango device and bone-conducting headphones the system is based on (left) and both in use in an experiment (right).}\label{fig:tango-headphone}
\end{figure}

@@

\section{Interface Description}

\subsection{Hardware Selection}

Electronic navigation aids and guidance systems for people with vision impairments have had trouble appealing to the visually impaired community at large because of problems that are typical of academic research and prototypes: prohibitive costs and user unfriendly interfaces and hardware requirements, e.g.\ head-mounted cameras and GPS antennae~\cite{golledge2004stated,yusif2016older,arditi2013user}.
Indeed, the traditional walking cane remains the go-to tool for navigation and obstacle avoidance. 
With the ActiVis system, we hope to tackle the issue of cost, user-acceptance and usability by using as many as possible `standard' devices and hardware as possible, with the idea that no specialist hardware is required and the user will not stand out if seen walking in public with a phone in their hand and wearing a headset.
The ActiVis system is based on a concept proposed in~\cite{bellotto2013multimodal,lock2017portable}, which uses a Google Tango device that is an Android-based device that comes equipped with an RGB-D camera to sense colour and depth and combines an inertial measurement unit (IMU) with powerful and robust landmark recognition and image processing algorithms to localise itself in real-time.
An benefit of such a platform is its familiar, compact form-factor which will help overcome the hurdle of user-acceptance and usability.

A set of bone-conducting headphones was selected as the audio transmission medium.
These headphones sit on a user's cheekbones and conduct the audio signals through the skull into the inner ear, instead of through the outer ear like typical over-ear headphones do. 
This has the benefit of not denying the user access to ambient sounds and noise, which could prove dangerous to a person with limited vision that relies on sound to detect oncoming vehicles and people, for example~\cite{lichtenstein2012headphone}.
Open-back headphones that allow ambient noise through were also considered, but the AfterShockz headphones pictured in \cref{fig:tango-headphone} were ultimately selected since they do not interfere at all with incoming sound (open-back headphones are better than closed back ones in this regard, but still muffle incoming sound) and are also more discreet than the larger over-ear headphone models.

@@

\begin{figure}
  \centering
  \subfloat[]{\label{fig:cam-coords}\includegraphics[width=0.3\columnwidth]{figures/camera_coordinate.pdf}}
~
  \subfloat[]{\label{fig:pitch-gain}\includegraphics[clip, trim=0 0 0 40, width=0.6\columnwidth]{figures/pitch_gain_functions.png}}
  \caption{The reference system used by the guidance interface showing the camera vector and pan and elevation angles (left) and the pitch gain function used to convey the target's elevation angle (right). }
\end{figure}

@@

The human audition system uses binaural comparison cues, such as the inter-aural time and level differences (ITD and ILD respectively), to localise a sound source on the horizontal plane~\cite{blauert1969sound}.
The ITD is the perceived time delay between the signal reaching both ears, while the ILD is the perceived volume difference in the signal.
For example, a sound that comes from the individual's right, will hit the right ear first with a slightly higher volume.

In this work, a pure sinusoidal wave refreshed at \SI{10}{\hertz} was used. 
People typically have trouble localising a pure tone devoid of a rich spectral signature.
However, the ITD and ILD is independent of the tone's spectral make-up and the elevation angle is given through a different mechanism.
A pure sine wave is therefore sufficient to convey the target's pan angle.
To transform and spatialise the audio signal, we used OpenAL's default HRTF, based on the MIT's KEMAR dataset~\cite{hiebert2005openal,gardner1995hrtf}, that uses the user and targets' positions as input and outputs a transformed audio signal.

@@

Using a generic HRTF, such as OpenAL's KEMAR implementation, with bone-conducting headphones has been shown to have trouble conveying the elevation angle of a sound source accurately~\cite{macdonald2006spatial,schonstein2008comparison}.
To compensate for this, we convey the target's elevation angle by adjusting the tone's pitch (i.e.\ the sine wave's frequency) as a function of the elevation. 
When the camera vector is at the target's elevation, the tone is set to neutral pitch.
Then, when the target is above the camera vector, the pitch is increased and decreased when the target is below the camera vector.
This high/low association scheme is used given humans' natural association of high-pitched sounds with elevated sound sources and lower-pitched sounds with source's below the individual's earline~\cite{pratt1930spatial,blauert1997spatial}.
A logarithmic, octave and semi-tone based function is used to adjust the tone's pitch to ensure perceptible changes while keeping the timbre roughly constant~\cite{shepard1964circularity}.
The pitch is updated at a rate of \SI{10}{\hertz} and changes as the user moves the device.

The pitch is changed as a linear function of the elevation angle and the gradient is determined by setting the angle and pitch limits.
For this work we only consider a \SI{180}{\degree} field of view in front of the user and limit the elevation angle to a range of $[-\frac{\pi}{2}, \frac{\pi}{2}]$.
The pitch limits are set at some integer number of octaves above and below the neutral, on-elevation pitch.
After practical tests with the interface, we set the neutral pitch to \SI{512}{\hertz}, which is comfortably audible and allows for a large number of audible octave limits to be selected.
For this work, we set the pitch limits to 2 octaves away from the neutral pitch, giving frequency limits of [\SI{128}{\hertz}, \SI{2048}{\hertz}].
The linear function is visualised in \cref{fig:pitch-gain}.
